The Evolution of Phenotypically Invariant Genetic Networks

### 2017-11-28 kalman_walk_paper.tex

Consistent with empirical observation \js{cite}, we see that the fitnesses of $F_2$ hybrids drop at a much faster rate than those of $F_1$s,
since $F_1$ and $F_2$ phenotypes diverge linearly and at the square root of time relative to parentals, respectively.
\plr{refer to Turelli here}
% In this model, this is due to recombination shuffling system architectures, leaving the $F_2$ system homozygous for one parent at one particular locus and homozygous for the other parent at another possibly incompatible locus.
Another natural consequence of the model is Haldane's rule,
that if only one $F_1$ hybrid sex is inviable or sterile it is likely to be the heterogametic sex.
This occurs because if the genes underlying a regulatory network are distributed among both autosomes and the sex chromosome, 
then heterogametic $F_1$s show variation similar to that seen in $F_2$s.

Is there evidence that this is actually occurring?
System drift and network rewiring has been inferred across the tree of life \citep{dalal2017transcription, johnson2017rewiring},
and there is often significant regulatory variation segregating within populations \citep{XXX}.
\js{cite sergey?}
Transcription in hybrids between closely related species with conserved transcriptional patterns can also be divergent 
\citep{haerty2006gene, maheshwari2012cis, coolon2014tempo, michalak2004association}.
Furthermore, in cryptic species complexes (e.g., sun skinks \citep{barley2013challenge}),
strongly reproductively incompatible species may be nearly morphologically indistinguishable.
\plr{check the skinks have RI!!}
%Of course, selective pressures have likely not been constant over evolutionary time
%in any of these examples,
%but the process we describe here will still occur.
%This model therefore may provide a neutral ``null'' model of gene network evolution, a strategy advocated for by \citet{koonin2016splendor}.
\plr{revisit}
Of course the basic assumptions of this model will be violated in practice (constant selective pressures, etc.), however, this model can function as a ``neutral null'' description of gene network evolution (a strategy advocated for in \citet{lynch2007frailty, fay2008evaluating, koonin2016splendor}). If this model succeeds in describing actual system evolution, it can be inferred that the mechanism underlying species formation does not require the inclusion of adaptation or changes in selection to explain the emergence of hybrid incompatibility.

@@

or a conseuqence of genetic conflict \citep{examples}.

>>>>

Consistent with empirical observation of \emph{hybrid breakdown} \citep[e.g.][]{plotner2017chlorosis}), we see that the fitnesses of $F_2$ hybrids drop at a much faster rate than those of $F_1$s,
since $F_1$ and $F_2$ phenotypes diverge linearly and at the square root of time relative to parentals, respectively.
\plr{refer to Turelli here}
% In this model, this is due to recombination shuffling system architectures, leaving the $F_2$ system homozygous for one parent at one particular locus and homozygous for the other parent at another possibly incompatible locus.
Another natural consequence of the model is Haldane's rule,
that if only one $F_1$ hybrid sex is inviable or sterile it is likely to be the heterogametic sex.
This occurs because if the genes underlying a regulatory network are distributed among both autosomes and the sex chromosome, 
then heterogametic $F_1$s show variation similar to that seen in $F_2$s.

Is there evidence that this is actually occurring?
System drift and network rewiring has been inferred across the tree of life \citep{dalal2017transcription, johnson2017rewiring},
and there is often significant regulatory variation segregating within populations \citep{XXX}.
\js{cite sergey?}
Transcription in hybrids between closely related species with conserved transcriptional patterns can also be divergent 
\citep{haerty2006gene, maheshwari2012cis, coolon2014tempo, michalak2004association}.
Furthermore, in cryptic species complexes (e.g., sun skinks \citep{barley2013challenge}),
%strongly reproductively incompatible species
genetically distinct species
may be nearly morphologically indistinguishable.
\plr{check the skinks have RI!!} \jss{the paper just says highly ``genetically distinct''; nothing is stated about RI}
%Of course, selective pressures have likely not been constant over evolutionary time
%in any of these examples,
%but the process we describe here will still occur.
%This model therefore may provide a neutral ``null'' model of gene network evolution, a strategy advocated for by \citet{koonin2016splendor}.
\plr{revisit}
Of course the basic assumptions of this model will be violated in practice (constant selective pressures, etc.), however, this model can function as a ``neutral null'' description of gene network evolution (a strategy advocated for in \citet{lynch2007frailty, fay2008evaluating, koonin2016splendor}). If this model succeeds in describing actual system evolution, it can be inferred that the mechanism underlying species formation does not require the inclusion of adaptation or changes in selection to explain the emergence of hybrid incompatibility.

@@

or a conseuqence of genetic conflict \citep{presgraves2010molecular, crespi2013conflictual}.
